---
title: "Fire hazard for `r stringr::str_to_title(params$hvra_name)` in the `r stringr::str_to_title(params$landscape_name)` Landscape"
author: "Sophie Gilbert, Mike Koontz, Katharyn Duffy, Vibrant Planet"
date: "`r Sys.Date()`"
output: html_document 

  # params:
#   landscape_name: ""
#   hvra_vector_fname: ""
#   hvra_name: ""
#   flp_fname: ""
#   outyear_bp_fname: ""
#   aoi_fname: ""
#   rf: ""
---

```{r load libraries, echo=FALSE, message=FALSE, results=FALSE}
library(aws.s3)
library(dplyr)
library(kableExtra)
library(tmaptools)
library(colorRamps)
library(viridis)
library(tmap)
library(terra)
options(scipen = 999)

```

```{r fetch the aoi, echo=FALSE, message=FALSE}
aoi = sf::st_read(params$aoi_fname)
```

```{r get flp and bp data and crop, echo=FALSE, message=FALSE, results=FALSE, cache=TRUE}
flp = terra::rast(params$flp_fname)
outyear_bp = terra::rast(params$outyear_bp_fname)
```

# Map the HVRA footprint

The map below shows the location of the HVRA relative to the `params$landscape_name` landscape (or area of interest).

```{r fetch_hvra_footprint, echo=FALSE, message=FALSE}

if(params$hvra_vector_fname != "" & tools::file_ext(x = params$hvra_vector_fname) == "gpkg") {
  
  hvra_footprint_sf = sf::st_read(params$hvra_vector_fname) |> 
    sf::st_transform(sf::st_crs(flp))
  
} else {
  message("Must provide a path to a .gpkg representing the HVRA geometry.")
}

aoi.plot = sf::st_transform(sf::st_geometry(aoi), crs = sf::st_crs(hvra_footprint_sf))

plot(sf::st_transform(sf::st_geometry(aoi), crs = sf::st_crs(hvra_footprint_sf)))
plot(sf::st_geometry(hvra_footprint_sf), add = TRUE, col = "blue")
plot(
  sf::st_transform(
    sf::st_geometry(aoi), 
    crs = sf::st_crs(hvra_footprint_sf
    )
  ), 
  add = TRUE,
  border = NA,
  bg="transparent"
)

```

\newpage

# Predicted Fire Intensity

Here is a plot of the flame length probability raster, binned into flame length categories, across the area of interest. This is the fire behavior we would expect to see, if a given location were to burn, but does not reflect the probability of that location actually burning (see burn probability below).

```{r plot_flame_length_probability, echo=FALSE, warning = FALSE, message = FALSE, cache=TRUE}

#terra::plot(flp, col = viridis::cividis(100))

tmap_options(max.raster = c(plot = 2000, view = 2000))

tm_shape(flp)+
  tm_raster(palette=viridis::cividis(100),
            style = "cont",
            max.value = 1,
            title = "Probability",
            legend.revers = TRUE
  )+
  tm_layout(legend.outside = TRUE,
            panel.labels = c("0-2 ft. Flames", "2-4 ft. Flames", "4-6 ft. Flames",
                             "6-8 ft. Flames", "8-12 ft. Flames", ">12 ft. Flames"))

```

# 10-year cumulative burn probability

The map and histogram below demonstrate the 10-year cumulative burn probability for the area of interest, regardless of fire intensity.

> *For example, the number 0.25 would represent a 25% probability of that location (pixel) burning within the next 10 years, but does not reflect fire behavior (see above for probable flame lengths).*

```{r plot_burn_probability, echo=FALSE, message=FALSE, results=FALSE, fig.show="hold", out.width="50%", cache=TRUE}
#terra::plot(outyear_bp, col = viridis::inferno(100))

max.bp = max(values(outyear_bp), na.rm=T)
max.bp.round = round(max.bp, digits =1)

breaks.bp = seq(0, max.bp.round, by = 0.05)

tmap_options(max.raster = c(plot = 100000, view = 100000))

map_raster_bp <- tm_shape(outyear_bp, raster.downsample=T)+
  tm_raster(palette=viridis::inferno(100),
            style = "cont",
            max.value = 1,
            breaks= breaks.bp,
            title = "Probability",
            legend.reverse = T
  )+
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right",
            legend.format = list(fun = function(x) {
              ifelse(x %in% breaks.bp, x, "")
            }))

map_bp <- map_raster_bp +
  tm_shape(aoi.plot)+
  tm_borders(col="black")

map_bp

bp_hist <- hist(values(outyear_bp)[!is.na(values(outyear_bp))], breaks=breaks.bp, xlab="10 Year Cumulative Burn Probability", main = "Histogram of values", col=viridis::inferno(n = length(breaks.bp)))


```

```{r binned burn probability of AOI, echo=FALSE, results = 'asis', cache=TRUE}
#A single 30m pixel (on each side) = 900 Square Meters = 0.2223945 acres (1 square meter = 0.000247105 acres).

rr = paste(breaks.bp[1:length(breaks.bp)-1], breaks.bp[2:length(breaks.bp)], sep= "-")
bp_area = data.frame(Risk_Range = rr, Acres = round(bp_hist$count*0.2223945, digits = 0), Percent = round(bp_hist$count/sum(bp_hist$count), digits=4)*100)

bp_area  %>%
  kbl(caption = "Predicted wildfire exposure for the AOI, base on 10-year burn probability and broken into risk ranges, and expressed as acres and as a proportion of the total",  
      col.names = c("Range", "Acres", "Percent of Acres (%)"), format="pandoc", booktabs=T, format.args = list(big.mark = ","))


```

```{r calculate mean fire hazard, echo=FALSE, message=FALSE, cache=TRUE}
# Calculate fire hazard as the burn probability times the probabilistic flame
# length times the response function for the hvra for each flame length, then
# summing across those probabilistic effects per pixel to produce a single layer
# of hazard
fire_hazard = sum(outyear_bp * flp * params$rf)

# Crop the hazard layer to the hvra footprint and mask to the hvra footprint too
fire_hazard_cropped = terra::crop(
  x = fire_hazard, 
  y = hvra_footprint_sf, 
  mask = TRUE
)
```

```{r calculate burn probability of hvra, echo=FALSE, cache=TRUE}
#Crop the burn probability layer to the hvra footprint and mask to the hvra footprint too
outyear_bp_cropped = terra::crop(
  x = outyear_bp, 
  y = hvra_footprint_sf, 
  mask = TRUE
)

max.bp.hvra = round(max(values(outyear_bp_cropped), na.rm=T), digits=2)
min.bp.hvra = round(min(values(outyear_bp_cropped), na.rm=T), digits=2)


```

```{r calculate mean burn probability of hvra, echo=FALSE, cache=TRUE}
## Calculate the expected mean 10-year burn probability across the hvra footprint (expressed as a proportion)
hvra_10yr_burn_prob = exactextractr::exact_extract(
  x = outyear_bp_cropped, 
  y = hvra_footprint_sf, 
  fun = "mean"
)

```

## Analyze the 10-year burn probability of the HVRA

The map, histogram, and table below summarize the 10-year cumulative burn probability within the footprint of the `r hvra_name` HVRA of interest, which ranges from `r min.bp.hvra*100`% to `r max.bp.hvra*100`%, with a mean value of `r round(hvra_10yr_burn_prob, digits = 2)*100`%.

```{r plot burn probability of hvra, echo=FALSE, message = F, fig.show="hold", out.width="50%", cache=TRUE}
tmap_options(max.raster = c(plot = 100000, view = 100000))

map_raster_bp_cropped <- tm_shape(outyear_bp_cropped, raster.downsample=T)+
  tm_raster(palette=viridis::inferno(100),
            style = "cont",
            max.value = 1,
            breaks= breaks.bp,
            title = "Probability",
            legend.reverse = T
  )+
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right",
            legend.format = list(fun = function(x) {
              ifelse(x %in% breaks.bp, x, "")
            }))

map_bp_cropped <- map_raster_bp_cropped +
  tm_shape(aoi.plot)+
  tm_borders(col="black")


map_bp_cropped

bp_cropped_hist <- hist(values(outyear_bp_cropped)[!is.na(values(outyear_bp_cropped))], breaks = seq(0, max.bp.round, by = 0.05), xlab="10 Year Cumulative Burn Probability, hvra", main = "Histogram of values", col=viridis::inferno(n = length(breaks.bp)))


```

```{r binned burn probability of hvra, echo=FALSE, results = 'asis', cache=TRUE}
#A single 30m pixel (on each side) = 900 Square Meters = 0.2223945 acres (1 square meter = 0.000247105 acres).

rr = paste(breaks.bp[1:length(breaks.bp)-1], breaks.bp[2:length(breaks.bp)], sep= "-")
bp_area_cropped = data.frame(Risk_Range = rr, hvra_Acres = round(bp_cropped_hist$count*0.2223945, digits = 0), hvra_Percent = round(bp_cropped_hist$count/sum(bp_cropped_hist$count), digits=4)*100)

bp_area_cropped  %>%
  kbl(caption = "Predicted wildfire exposure for the HVRA, base on 10-year burn probability.  Risk is broken into ranges, and expressed both via acres and as a proportion",  
      col.names = c("Range", "Acres", "Percent of Acres (%)"), format="pandoc", booktabs=T, format.args = list(big.mark = ","))


```

## Analysis of expected fire behavior, expressed as flame length probabilities, within the HVRA

Flame length categories represent:

-   0-2 foot flame lengths
-   2-4 foot flame lengths
-   4-6 foot flame lengths
-   6-8 foot flame lengths
-   8-12 foot flame lengths
-   Greater than 12 foot flame lengths

```{r plot flame length in hvra, echo=FALSE, cache=TRUE}
#Crop the burn probability layer to the hvra footprint and mask to the hvra footprint too
flp_cropped = terra::crop(
  x = flp, 
  y = hvra_footprint_sf, 
  mask = TRUE
)


tmap_options(max.raster = c(plot = 100000, view = 100000))

tm_shape(flp_cropped)+
  tm_raster(palette=viridis::cividis(100),
            style = "cont",
            max.value = 1,
            title = "Probability",
            legend.revers = T
  )+
  tm_facets(ncol = 3)+
  tm_layout(legend.outside = TRUE,
            panel.labels = c("0-2 ft. Flames", "2-4 ft. Flames", "4-6 ft. Flames",
                             "6-8 ft. Flames", "8-12 ft. Flames", ">12 ft. Flames"))
```

```{r calculate flame lentgths of hvra, echo=FALSE, cache=TRUE}
## Calculate the expected mean probability of different flame lengths across the hvra footprint
hvra_flp_prob = exactextractr::exact_extract(
  x = flp_cropped, 
  y = hvra_footprint_sf, 
  fun = "mean"
)

#hvra_flp_prob2 = terra::extract(flp_cropped, hvra_footprint_sf, fun="mean")

```

```{r calculate expected change in hvra value, echo=FALSE}
## Calculate the expected mean change in hvra value across the whole landscape
hvra_value_change = exactextractr::exact_extract(
  x = fire_hazard_cropped, 
  y = hvra_footprint_sf, 
  fun = "mean"
)
```

```{r calculate conditional change in hvra value if it burns, echo=FALSE, cache=TRUE}
## Calculate the potential/conditional change in hvra value, if it burns
## Then calculate expected mean conditional change in hvra value across the whole landscape

conditional_loss = sum(flp * params$rf)
conditional_loss_all_flips = (hvra_flp_prob*params$rf)

# Crop the hazard layer to the hvra footprint and mask to the hvra footprint too
conditional_loss_cropped = terra::crop(
  x = conditional_loss, 
  y = hvra_footprint_sf, 
  mask = TRUE
)

hvra_potential_value_change = exactextractr::exact_extract(
  x = conditional_loss_cropped, 
  y = hvra_footprint_sf, 
  fun = "mean"
)
```

## Analysis of the potential change in `r params$hvra_name` HVRA value if a fire occurs

The map and histogram below demonstrate the expected response of the HVRA to predicted flame lengths, should a fire occur. Change in value represents a proportional change *(not percent change)*.

```{r Plot potential hvra change, echo=FALSE, message=FALSE, results=FALSE, fig.show="hold", out.width="50%"}

max.cl = max(values(conditional_loss_cropped), na.rm=T)
max.cl.round = round(max.cl, digits =1)

min.cl = min(values(conditional_loss_cropped), na.rm=T)
min.cl.round = round(min.cl, digits =1)

#breaks.cl = seq(min.cl.round, max.cl.round, by = 0.05)
breaks.cl = seq(-1, 1, by = 0.1)
col.cl = get_brewer_pal("RdYlBu", n=length(breaks.cl), plot=F)
tmap_options(max.raster = c(plot = 100000, view = 100000))

map_raster_cl <- tm_shape(conditional_loss_cropped, raster.downsample=T)+
  tm_raster(palette=get_brewer_pal("RdYlBu", n=100, plot=F),
            style = "cont",
            max.value = 1,
            breaks= breaks.cl,
            title = "Proportional Change",
            legend.reverse = T
  )+
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right",
            legend.format = list(fun = function(x) {
              ifelse(x %in% breaks.cl, x, "")
            }))

map_cl <- map_raster_cl +
  tm_shape(aoi.plot)+
  tm_borders(col="black")

map_cl

cl_hist_cropped <- hist(values(conditional_loss_cropped)[!is.na(values(conditional_loss_cropped))], breaks=breaks.cl, xlab="Potential proportional Change in hvra Value", main = "Histogram of values", col=col.cl)

```

```{r binned potential change (susceptibility), echo=FALSE, results = 'asis'}
#A single 30m pixel (on each side) = 900 Square Meters = 0.2223945 acres (1 square meter = 0.000247105 acres).

rfr = paste(breaks.cl[1:length(breaks.cl)-1], breaks.cl[2:length(breaks.cl)], sep= "-")
cond_loss_area = data.frame(RF_Range = rfr, Acres = round(cl_hist_cropped$count*0.2223945, digits = 0), Percent = round(cl_hist_cropped$count/sum(cl_hist_cropped$count), digits=4)*100)

cond_loss_area  %>%
  kbl(caption = "Predicted proportional change in hvra value, should a fire occur (conditional loss/gain)",  
      col.names = c("Range", "Acres", "Percent of Acres (%)"), format="pandoc", booktabs=T, format.args = list(big.mark = ","))


```


## Analysis of the expected change in `r params$hvra_name` value due to wildfire hazard under a "no action" scenario

This fire hazard analysis includes:

  a)  expected fire behavior (flame lengths) if a burn occurs,
  
  b)  10-year probability of a burn occurring, and
  
  c)  expected `r params$hvra_name` response to fire behavior (response functions) to get to an expected 10-year change in value.
  

```{r plot fire hazard of hvra, echo=FALSE, message=FALSE, results=FALSE, fig.show="hold", out.width="50%"}

max.fh = max(values(fire_hazard_cropped), na.rm=T)
max.fh.round = round(max.fh, digits =1)

min.fh = min(values(fire_hazard_cropped), na.rm=T)
min.fh.round = round(min.fh, digits =1)

#breaks.cl = seq(min.cl.round, max.cl.round, by = 0.05)
breaks.fh = seq(-1, 1, by = 0.1)
col.fh = get_brewer_pal("RdYlBu", n=length(breaks.cl), plot=F)
tmap_options(max.raster = c(plot = 100000, view = 100000))

map_raster_fh <- tm_shape(fire_hazard_cropped, raster.downsample=T)+
  tm_raster(palette=get_brewer_pal("RdYlBu", n=100, plot=F),
            style = "cont",
            max.value = 1,
            breaks= breaks.cl,
            title = "Proportional Change",
            legend.reverse = T
  )+
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right",
            legend.format = list(fun = function(x) {
              ifelse(x %in% breaks.fh, x, "")
            }))

map_fh <- map_raster_fh +
  tm_shape(aoi.plot)+
  tm_borders(col="black")

map_fh

fh_hist_cropped <- hist(values(fire_hazard_cropped)[!is.na(values(fire_hazard_cropped))], breaks=breaks.fh, xlab="Expected proportional change in HVRA Value", main = "Histogram of values", col=col.cl)
```

```{r binned expected change (susceptibility), echo=FALSE, results = 'asis'}
#A single 30m pixel (on each side) = 900 Square Meters = 0.2223945 acres (1 square meter = 0.000247105 acres).

fhr = paste(breaks.fh[1:length(breaks.fh)-1], breaks.fh[2:length(breaks.fh)], sep= "-")
cond_loss_area = data.frame(RF_Range = rfr, Acres = round(fh_hist_cropped$count*0.2223945, digits = 0), Percent = round(fh_hist_cropped$count/sum(fh_hist_cropped$count), digits=4)*100)

cond_loss_area  %>%
  kbl(caption = "Predicted Proportional change in HVRA value: No Action Scenario",  
      col.names = c("Range", "Acres", "Percent of Acres (%)"), format="pandoc", booktabs=T, format.args = list(big.mark = ","))


```

## Summary

### Across the `r params$landscape_name` landscape, we expect the `r params$hvra_name` HVRA to `r ifelse(test = hvra_value_change > 0, yes = "increase in", no = ifelse(test = hvra_value_change == 0, yes = "stay the same", no = "decrease in"))` value over the next 10 years.

### The 10-year average burn probability is `r round(hvra_10yr_burn_prob, digits = 3)*100`%.

### The potential change in value if it burns, on average, would be `r round(hvra_potential_value_change, digits = 3)*100` %.

### The response of `r params$hvra_name` to fire hazard (which factors in the 10-year burn probability, the expected flame lengths if it burns, and the expected species response to those flame lengths) under 'No-Action' will be `r round(hvra_value_change, digits = 3)*100` % change in value.

```{r SummaryTable, echo=FALSE, message = FALSE, warning = FALSE, results = 'asis'}
output = cbind(c("Mean probability (%)", "RF (% change)", "Potential change (%)"), rbind(round(hvra_flp_prob, digits=2)*100, round(params$rf, digits=2)*100, round(conditional_loss_all_flips, digits=2)*100))
colnames(output) <- c("Variable", "0-2 ft", "2-4 ft", "4-6 ft", "6-8 ft", "8-12 ft", ">12 ft")


output  %>%
  kbl(caption = "Expected flame length probability and HVRA expected response (% change in value)",  format="pandoc", booktabs=T)


```
